---
output:
  pdf_document
geometry: margin=1.0in
fontsize: 11pt
numbersections: false
bibliography: reference.bib
csl: evolution-and-human-behavior.csl
header-includes:
  - \usepackage{lscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
  - \newcommand{\bcenter}{\begin{center}}
  - \newcommand{\ecenter}{\end{center}}
  - \newcommand{\btitlepage}{\begin{titlepage}}
  - \newcommand{\etitlepage}{\end{titlepage}}
  - \usepackage{setspace}\doublespacing
  - \usepackage{helvet}
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{booktabs}
  - \usepackage[font=small,labelfont=bf]{caption}
  - \pagenumbering{gobble}
---

```{r set_up, include=FALSE}
library(tidyverse); library(here); library(knitr); library(kableExtra)

options(knitr.kable.NA = '')

# main analyses results
results_inpool <-
  here("data/processed/des_and_test_result.csv") %>%
  read_csv() %>%
  rename(hls = homosexual_dlikelihood_dstandard_lt) %>%
  mutate(across(.col = c(f_inpool.mean:m_outpool.sd,
                inpool.conf.low:inpool.conf.high,
                outpool.conf.low:outpool.conf.high),
                ~ .x/150)) %>%
  select(hls, m_inpool.mean, m_inpool.sd, f_inpool.mean, f_inpool.sd) %>% 
  filter(str_detect(hls, "^0"))

# sensitivity analyses results
sa_results <-
  here("data/processed/sa_des_and_test_result.csv") %>%
  read_csv() %>%
  rename(hls = homosexual_dlikelihood_dstandard_lt) %>%
  mutate(across(.col = c(f_inpool.mean:m_outpool.sd,
                inpool.conf.low:inpool.conf.high,
                outpool.conf.low:outpool.conf.high),
                ~ .x/150)) %>%
  mutate(across(contains("p.value"), ~format(.x, digits = 2, nsmall = 3))) %>%
  mutate(across(contains("p.value"), ~ ifelse(.x >= .001,
                                              str_remove(.x, "^0"),
                                              "< .001")
                                              ))

sa_results_hg <-
  here("data/processed/sa_compare_hetero_gay_men.csv") %>%
  read_csv() %>%
  rename(ls = dlikelihood_dstandard_lt) %>%
  mutate(across(contains("p.value"), ~format(.x, digits = 2, nsmall = 3))) %>%
  mutate(across(contains("p.value"), ~ ifelse(.x >= .001,
                                              str_remove(.x, "^0"),
                                              "< .001")
                                              ))

# define inline hook
inline_hook <- function(x) {

    if (is.numeric(x)) {

      formatted <- ifelse(
        abs(x - round(x)) < .Machine$double.eps & x != 0,
        # treat as integer
        format(x, digits=0, big.mark = ","),
        # treat as floating point number
        format(x, digits=1, nsmall=2, big.mark = ",")
        ) %>%
        str_replace("-", "–")

    } else {

      formatted <- x

    }

}

knit_hooks$set(inline = inline_hook)
```

\btitlepage

\bcenter

\vspace*{30mm}

# Supplemental Materials to "An Agent-Based Model of Sex and Sexual Orientation Differences in Short-Term Mating Behaviors as a Result of Mating Preferences"

\ecenter

\etitlepage

\newpage

\bcenter

## A. Overview, Design Concepts, and Details of the Model

\ecenter

This description of the model follows the standard ODD protocol of reporting individual-based and agent-based modeling [@grimm_standard_2006].

### Purpose

The model simulated sex differences in short-term mating behavior. Findings in evolutionary psychology suggest that males have a stronger desire for short-term mating than females [@regan_minimum_1998]. They are also less choosy regarding short-term mating partners [@buss2019mate; @buss_sexual_1993]. This model explores both heterosexual and gay males’ as well as heterosexual and lesbian females' behavioral outcomes under these assumptions.

The first version of the model simulated short-term mating behaviors among heterosexual individuals, short-term mating and/or long-term relationships were formed only between individuals of the opposite sex. The second version of the model simulated the behaviors among gay males and lesbian females, where individuals only had short-term matings and/or formed long-term relationships with those of the same sex. The key parameters were individuals’ likelihood of engaging in short-term mating and their standards for a short-term partner. We varied those parameters to examine conditions under which sex differences in behavioral outcomes may emerge.

### Entities, state variables, and scales

The model included two types of entities: global and agents. In this model, time and space were modeled as discrete variables. During each time step, agents executed the commands described in the schedule. Space was represented as discrete locations on a two-dimensional 33*33 lattice. All state variables are summarized in Supplemental Table 1.

### Process overview and scheduling

The model was written using Netlogo 6.2.1 [@wilensky_1999]. The model proceeded in discrete time steps, and the entities executed commands in the following order:

\begin{enumerate}

  \item Setting up (Global, only executed once)
  \begin{enumerate}
    \item 300 agents were created with all state variables set to the default
    \item A number of agents were set to either males (number-of-male) or females (300 – number-of-men) with their respective short-term mating likelihood (short-term-likelihood), and standard for a short-term mate (mating-standard) 
  \end{enumerate}
  
  
  \item Moving (The following procedures were executed by agents unless specified otherwise)
  
  \begin{enumerate}
    \item If not coupled
    \begin{enumerate}
      \item If staying within a distance of movement-range from the initial point on both the x and y axis, randomly set their heading and move by a random distance less than movement-range / 2
      \item If staying outside the distance of movement-range from the initial point on either the x or y axis, head toward the initial point and move by a random distance less than movement-range / 2
    \end{enumerate}
    \item If coupled, do not move
  \end{enumerate}
  
  
  \item Deciding on short-term mating behavior
  \begin{enumerate}
    \item Set short-term-mate? to true by short-term-likelihood
  \end{enumerate}
  
  
  \item Short-term and long-term mating (To avoid double count, only females execute this chunk. This does not affect the model outcomes.)
  \begin{enumerate}
    \item If any males here, females randomly select one as the target for short-term mating
    \begin{enumerate}
      \item If the target has desirability higher than her mating-standard \& she has desirability higher than the target’s mating-standard \& both of them have short-term-mate? set to true,
      \begin{enumerate}
        \item Both of them add 1 to short-term-count
        \item Both of them add each other’s ID number to their short-term-history if not in it yet
      \end{enumerate}
    \end{enumerate}
    \item If any males here and females are not coupled, females randomly select one as the potential partner for long-term mating
    \begin{enumerate}
      \item By long-term-likelihood, they form a long-term relationship, and
      \begin{enumerate}
        \item Both of them set coupled? to true
        \item Both of them set each other as their long-term partner
      \end{enumerate}
    \end{enumerate}
  \end{enumerate}
  
  \item Updating state variables
  \begin{enumerate}
    \item Set number-of-partner-short the length of short-term-history
    \item If coupled? is true and short-term-likelihood equals male-/female-short-term-likelihood, set short-term-likelihood to a third of its value
  \end{enumerate}
  
\end{enumerate}

### Design concepts

*Sensing.* Agents were able to detect other agents’ mate value so that they could decide whether to have short-term mating with them.

*Interaction.* Females and males interacted to have short-term mating behaviors or form long-term relationships.

*Stochasticity.* Agents' movement in this model had stochastic elements, with the heading and the movement distance being randomly selected. Whether agents decided to have short-term mating or form long-term relationships was also randomly determined.

*Observation.* Simulations were run for 1,000 time steps. The dependent variables were measured at the end of each run.

\newpage

\bcenter

## B. Additional Analyses

\ecenter

\textbf{Supplemental Table B1: Mean proportion of heterosexual males and females in the mating pool and its standard deviation after 1,000 time steps in the model.}

\begingroup
\renewcommand{\arraystretch}{1.5}

```{r inpool,echo=FALSE}
results_inpool %>%
  mutate(across(where(is.numeric), inline_hook)) %>%
  mutate(
    condition =
      c("Same likelihood*Same standard",
        "Same likelihood*Different standard",
        "Different likelihood*Same standard",
        "Different likelihood*Different standard")
  ) %>%
  select(-hls) %>% 
  column_to_rownames("condition") %>%
  kable(
    digits = 2, format = "latex", booktabs = TRUE, align = "cccc",
    row.names = TRUE, col.names = rep(c("M", "SD"), 2)
    ) %>%
  row_spec(0, italic = TRUE, align = "c") %>%
  add_header_above(
    c(" " = 1,
      "Males" = 2, "Females" = 2)
    ) %>%
  column_spec(1, width = "2.5in") %>%
  column_spec(c(2, 3, 4, 5), width = "0.8in") %>%
  footnote(
    general = "\\\\textit{Note.} \\\\textit{N} = 2,500 in each condition.",
    general_title = "", escape = FALSE,
    threeparttable = TRUE, fixed_small_size = TRUE) %>%
  kable_styling(position = "center", latex_options = "hold_position")
```

\endgroup

\newpage

\bcenter

## C. Sensitivity Analyses

\ecenter

\renewcommand{\tablename}{Supplemental Table}
\renewcommand{\figurename}{Supplemental Figure}

This section presents the results from sensitivity analyses to complement the main results presented in the paper. We set the parameter long-term-likelihood to 0, 2, 4, 6, 8, and repeated the same set of statistical analyses as the main study for each value. This was to show that the results were robust across different assumptions on the rate of forming long-term relationship. The simulations for sensitivity analyses adopted the same experimental designs but was run only 100 times for each set of parameter setting.

### Sensitivity Analysis on Sex Differences in Short-term Mating Behaviors in the Full Population

\newpage


\renewcommand{\thefigure}{C1}
\begin{figure}[h]
  \centering
  \includegraphics{figures/sa_fig1_men_vs_women.pdf}
  \caption{\textbf{Short-term mating behaviors of males and females for 0-8\% long-term relationship likelihood when sex differences existed in mating preferences.} Line plots summarizing the outcome variables separately for heterosexual individuals and gay males and lesbian females. Plot (A) shows the average number of short-term mating experiences, and plot (B) shows the average number of short-term mates. The points show mean values and whiskers represent standard errors, but the standard errors are small and overlap to form a single bar in some sets of parameter setting. The means were calculated using the full population of males and females in the model (\textit{N\textsubscript{male}} = \textit{N\textsubscript{female}} = 150).}
  \label{fig:men_vs_women}
\end{figure}


```{r table_1, include=FALSE, echo=FALSE}
# note the following code is tentative and the latex code of the table is mannually editted
sa_results %>%
  filter(str_detect(hls, pattern = "^(.)_1_1_(.)")) %>% 
  mutate(
    exp_all.d = abs(exp_all.d),
    partner_all.d = abs(partner_all.d)
  ) %>%
  mutate(across(where(is.numeric), inline_hook)) %>%
  unite(exp_all.conf.low, exp_all.conf.high, sep = ", ",
        col = "exp_all.conf") %>% 
  unite(partner_all.conf.low, partner_all.conf.high, sep = ", ",
        col = "partner_all.conf") %>% 
  mutate(
    condition = seq(1, 10, by = 1),
    exp_all.conf = paste0("[", exp_all.conf, "]"),
    partner_all.conf = paste0("[", partner_all.conf, "]")
    ) %>%
  select(condition, m_exp_all.mean, m_exp_all.sd,
         f_exp_all.mean, f_exp_all.sd, 
         exp_all.conf, exp_all.d,
         m_partner_all.mean, m_partner_all.sd,
         f_partner_all.mean, f_partner_all.sd, 
         partner_all.conf, partner_all.d)%>%
  column_to_rownames("condition") %>%
  kable(
    digits = 2, format = "latex", booktabs = TRUE,
    caption = "\\textbf{Short-term mating behaviors of men and women for 0-8 long-term relationship likelihood when gender differences existed in mating preferences}",
    row.names = TRUE, 
    col.names = c(rep(c(rep(c("M", "SD"), 2), "Difference CI [LL, UL]","d"), 2))
    ) %>%
  row_spec(0, italic = TRUE, align = "c") %>%
  add_header_above(
    c(" " = 1,
      "Men" = 2, "Women" = 2, " " = 2,
      "Men" = 2, "Women" = 2, " " = 2)
    ) %>%
  add_header_above(
    c(" " = 1,
      "Average short-term mating\nexperiences" = 6,
      "Average short-term\nmates" = 6)
    ) %>%
  column_spec(1, width = "0.5in") %>%
  footnote(
    general = "\\\\textit{Note.} The means were calculated using the full population of men and women in the model (\\\\textit{N\\\\textsubscript{men}} = \\\\textit{N\\\\textsubscript{women}} = 150).",
    general_title = "", escape = FALSE,
    threeparttable = TRUE, fixed_small_size = TRUE) %>% 
  kable_styling(latex_options = "scale_down")
```

\newpage

\blandscape

\begingroup
\renewcommand{\arraystretch}{1.5}

\renewcommand{\thetable}{C1}
\begin{table}

\caption{\textbf{Short-term mating behaviors of males and females for 0-8\% long-term relationship likelihood when sex differences existed in mating preferences}}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{0.5in}cccccccccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{6}{c}{Average short-term mating experiences} & \multicolumn{6}{c}{Average short-term mates} \\
\cmidrule(l{3pt}r{3pt}){2-7} \cmidrule(l{3pt}r{3pt}){8-13}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Males} & \multicolumn{2}{c}{Females} & \multicolumn{2}{c}{ } & \multicolumn{2}{c}{Males} & \multicolumn{2}{c}{Females} & \multicolumn{2}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11}
\multicolumn{1}{c}{\makecell[c]{Long-term\\Likelihood (\%)}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\makecell[c]{Mean difference\\95\%CI [LL, UL]}} & \multicolumn{1}{c}{\makecell[c]{Cohen's \\ \em{d}}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\makecell[c]{Mean difference\\95\%CI [LL, UL]}} & \multicolumn{1}{c}{\makecell[c]{Cohen's \\ \em{d}}}\\
\midrule
\multicolumn{13}{l}{\textit{Experiment 1: Heterosexual individuals}} \\
\multicolumn{1}{c}{0} & 5.94 & 0.61 & 5.94 & 0.61 & {}[–0.17, 0.17] & 0.00 & 3.43 & 0.34 & 3.43 & 0.34 & {}[–0.10, 0.10] & 0.00\\
\multicolumn{1}{c}{2} & 2.57 & 0.30 & 2.57 & 0.30 & {}[–0.08, 0.08] & 0.00 & 1.65 & 0.17 & 1.65 & 0.17 & {}[–0.05, 0.05] & 0.00\\
\multicolumn{1}{c}{4} & 1.97 & 0.29 & 1.97 & 0.29 & {}[–0.08, 0.08] & 0.00 & 1.23 & 0.15 & 1.23 & 0.15 & {}[–0.04, 0.04] & 0.00\\
\multicolumn{1}{c}{6} & 1.66 & 0.27 & 1.66 & 0.27 & {}[–0.07, 0.07] & 0.00 & 1.02 & 0.13 & 1.02 & 0.13 & {}[–0.04, 0.04] & 0.00\\
\multicolumn{1}{c}{8} & 1.53 & 0.26 & 1.53 & 0.26 & {}[–0.07, 0.07] & 0.00 & 0.88 & 0.12 & 0.88 & 0.12 & {}[–0.03, 0.03] & 0.00\\
\addlinespace
\multicolumn{13}{l}{\textit{Experiment 2: Gay males and lesbian females}} \\
\multicolumn{1}{c}{0} & 8.75 & 0.76 & 1.09 & 0.21 & {}[–7.81, –7.50] & 13.71 & 5.42 & 0.43 & 0.88 & 0.16 & {}[–4.64, –4.46] & 14.04\\
\multicolumn{1}{c}{2} & 5.02 & 0.48 & 0.61 & 0.16 & {}[–4.51, –4.31] & 12.45 & 3.38 & 0.30 & 0.51 & 0.13 & {}[–2.93, –2.80] & 12.20\\
\multicolumn{1}{c}{4} & 3.92 & 0.47 & 0.45 & 0.12 & {}[–3.56, –3.37] & 10.04 & 2.67 & 0.28 & 0.37 & 0.09 & {}[–2.35, –2.23] & 10.90\\
\multicolumn{1}{c}{6} & 3.22 & 0.43 & 0.40 & 0.11 & {}[–2.91, –2.73] & 8.95 & 2.21 & 0.26 & 0.33 & 0.08 & {}[–1.94, –1.83] & 9.95\\
\multicolumn{1}{c}{8} & 2.88 & 0.39 & 0.36 & 0.09 & {}[–2.60, –2.44] & 8.80 & 1.92 & 0.22 & 0.28 & 0.07 & {}[–1.68, –1.59] & 10.12\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item \textit{Note.} \textit{M} = Mean. \textit{SD} = Standard deviation. CI = Confidence interval. LL = Lower limit. UL = Upper limit. The means were calculated using the full population of males and females in the model (\textit{N\textsubscript{male}} = \textit{N\textsubscript{female}} = 150).
\end{tablenotes}
\end{threeparttable}
\end{table}

\endgroup

\elandscape

\newpage


### Sensitivity Analysis on Heterosexual Individuals in the Mating Pool

\renewcommand{\thefigure}{C2}
\begin{figure}[h]
  \centering
  \includegraphics{figures/sa_fig2_men_vs_women_inpool.pdf}
  \caption{\textbf{Short-term mating behaviors of heterosexual males and females in the mating pool for 0-8\% long-term relationship likelihood when sex differences existed in mating preferences.} Line plots summarizing the outcome variables. Plot (A) shows the average number of short-term mating experiences in the mating pool, and plot (B) shows the average number of short-term mates in the mating pool. Central points show mean values and whiskers represent standard errors, but the standard errors are small and overlap to form a single bar in some sets of parameter setting. The means were calculated using the population of heterosexual males and females who successfully participated in the mating pool (See Table C2 for the mean and standard deviation of the proportion of individuals in the mating pool).}
  \label{fig:men_vs_women_inpool}
\end{figure}

\newpage

\blandscape


```{r table_2, include=FALSE, echo=FALSE}
sa_results %>%
  filter(str_detect(hls, pattern = "^0_1_1")) %>%
  mutate(
    inpool.d = abs(inpool.d),
    exp_ip.d = abs(exp_ip.d),
    partner_ip.d = abs(partner_ip.d)
  ) %>%
  mutate(across(where(is.numeric), inline_hook)) %>%
  unite(inpool.conf.low, inpool.conf.high, sep = ", ",
        col = "inpool.conf") %>% 
  unite(exp_ip.conf.low, exp_ip.conf.high, sep = ", ",
        col = "exp_ip.conf") %>% 
  unite(partner_ip.conf.low, partner_ip.conf.high, sep = ", ",
        col = "partner_ip.conf") %>% 
  mutate(
    condition = seq(0, 8, by = 2),
    inpool.conf = paste0("[", inpool.conf, "]"),
    exp_ip.conf = paste0("[", exp_ip.conf, "]"),
    partner_ip.conf = paste0("[", partner_ip.conf, "]")
    ) %>%
  select(condition, 
         m_inpool.mean, m_inpool.sd,
         f_inpool.mean, f_inpool.sd,
         inpool.conf, inpool.d,
         m_exp_ip.mean, m_exp_ip.sd,
         f_exp_ip.mean, f_exp_ip.sd, 
         exp_ip.conf, exp_ip.d,
         m_partner_ip.mean, m_partner_ip.sd,
         f_partner_ip.mean, f_partner_ip.sd, 
         partner_ip.conf, partner_ip.d) %>%
  column_to_rownames("condition") %>%
  kable(
    digits = 2, format = "latex", booktabs = TRUE,
    caption = "\\textbf{Short-term mating behaviors of heterosexual men and women in the mating pool for 0-8 long-term relationship likelihood when gender differences existed in mating preferences}",
    row.names = TRUE, 
    col.names = c(rep(c(rep(c("M", "SD"), 2), "Mean difference 95 CI [LL, UL]","d"), 3))
    ) %>%
  row_spec(0, italic = TRUE, align = "c") %>%
  add_header_above(
    c(" " = 1,
      "Men" = 2, "Women" = 2, " " = 2,
      "Men" = 2, "Women" = 2, " " = 2,
      "Men" = 2, "Women" = 2, " " = 2)
    ) %>%
  add_header_above(
    c(" " = 1,
      "Mean proportion of individuals in the mating pool" = 6,
      "Average short-term mating experiences (in-pool)" = 6,
      "Average short-term mates (in-pool)" = 6)
    ) %>%
  column_spec(1, width = "0.5in") %>%
  footnote(
    general = "\\\\textit{Note.} \\\\textit{M} = Mean. \\\\textit{SD} = Standard deviation. CI = Confidence interval. LL = Lower limit. UL = Upper limit. The means were calculated with subsamples of men and women who had ever engaged in short-term mating behaviors.",
    general_title = "", escape = FALSE,
    threeparttable = TRUE, fixed_small_size = TRUE) %>% 
  kable_styling(latex_options = "scale_down")
```

\begingroup
\renewcommand{\arraystretch}{1.5}

\renewcommand{\thetable}{C2}
\begin{table}

\caption{\textbf{Proportion and short-term mating behaviors of heterosexual males and females in the mating pool for 0-8\% long-term relationship likelihood when sex differences existed in mating preferences}}
\centering
\begin{threeparttable}
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{0.3in}cccccccccccccccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{6}{c}{Mean proportion of individuals in the mating pool} & \multicolumn{6}{c}{Average short-term mating experiences (in-pool)} & \multicolumn{6}{c}{Average short-term mates (in-pool)} \\
\cmidrule(l{3pt}r{3pt}){2-7} \cmidrule(l{3pt}r{3pt}){8-13} \cmidrule(l{3pt}r{3pt}){14-19}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Males} & \multicolumn{2}{c}{Females} & \multicolumn{2}{c}{ } & \multicolumn{2}{c}{Males} & \multicolumn{2}{c}{Females} & \multicolumn{2}{c}{ } & \multicolumn{2}{c}{Males} & \multicolumn{2}{c}{Females} & \multicolumn{2}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11} \cmidrule(l{3pt}r{3pt}){14-15} \cmidrule(l{3pt}r{3pt}){16-17}
\multicolumn{1}{c}{\makecell[c]{Long-term\\Likelihood\\(\%)}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\makecell[c]{Mean difference\\95\%CI [LL, UL]}} & \multicolumn{1}{c}{\makecell[c]{Cohen's \\ \em{d}}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\makecell[c]{Mean difference\\95\%CI [LL, UL]}} & \multicolumn{1}{c}{\makecell[c]{Cohen's \\ \em{d}}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\makecell[c]{Mean difference\\95\%CI [LL, UL]}} & \multicolumn{1}{c}{\makecell[c]{Cohen's \\ \em{d}}}\\
\midrule
\multicolumn{1}{c}{0} & 0.50 & 0.04 & 0.89 & 0.03 & {}[0.38, 0.40] & 10.79 & 11.87 & 0.67 & 6.67 & 0.59 & {}[–5.37, –5.02] & 8.25 & 6.85 & 0.34 & 3.85 & 0.33 & {}[–3.10, –2.91] & 8.98\\
\multicolumn{1}{c}{2} & 0.47 & 0.04 & 0.75 & 0.04 & {}[0.26, 0.28] & 7.19 & 5.43 & 0.42 & 3.44 & 0.31 & {}[–2.09, –1.89] & 5.42 & 3.48 & 0.22 & 2.21 & 0.16 & {}[–1.33, –1.22] & 6.52\\
\multicolumn{1}{c}{4} & 0.44 & 0.04 & 0.65 & 0.05 & {}[0.20, 0.23] & 4.53 & 4.50 & 0.51 & 3.01 & 0.32 & {}[–1.61, –1.37] & 3.51 & 2.81 & 0.22 & 1.88 & 0.14 & {}[–0.98, –0.88] & 4.97\\
\multicolumn{1}{c}{6} & 0.42 & 0.04 & 0.59 & 0.05 & {}[0.16, 0.19] & 3.67 & 3.96 & 0.48 & 2.80 & 0.33 & {}[–1.28, –1.05] & 2.85 & 2.43 & 0.20 & 1.72 & 0.12 & {}[–0.76, –0.67] & 4.26\\
\multicolumn{1}{c}{8} & 0.40 & 0.04 & 0.54 & 0.05 & {}[0.14, 0.16] & 3.36 & 3.87 & 0.51 & 2.81 & 0.36 & {}[–1.19, –0.94] & 2.43 & 2.24 & 0.19 & 1.62 & 0.12 & {}[–0.66, –0.57] & 3.92\\
\bottomrule
\end{tabular}}
\begin{tablenotes}
\small
\item \textit{Note.} \textit{M} = Mean. \textit{SD} = Standard deviation. CI = Confidence interval. LL = Lower limit. UL = Upper limit. The means were calculated with \\ subsamples of males and females who had ever engaged in short-term mating behaviors.
\end{tablenotes}
\end{threeparttable}
\end{table}

\endgroup


\elandscape

\newpage

### Sensitivity Analysis on the Comparison Between Heterosexual and Gay Males

\renewcommand{\thefigure}{C3}
\begin{figure}[h]
  \centering
  \includegraphics{figures/sa_fig3_hetero_vs_gay_men.pdf}
  \caption{\textbf{Short-term mating behaviors of heterosexual males and gay males for 0-8\% long-term relationship likelihood when sex differences existed in mating preferences.} Line plots summarizing the outcome variables. Plot (A) shows the average number of short-term mating experiences, and plot (B) shows the average number of short-term mates. Central points show mean values and whiskers represent standard errors, but the standard errors are small and overlap to form a single bar in some sets of parameter setting. The means were calculated using the full population of heterosexual males and gay mlaes in the model (\textit{N\textsubscript{hetero}} = \textit{N\textsubscript{gay}} = 150).}
  \label{fig:hetero_vs_gay}
\end{figure}

\newpage

\blandscape

```{r table_3, include=FALSE, echo=FALSE}
sa_results_hg %>%
  filter(str_detect(ls, pattern = "^1_1")) %>%
  mutate(
    exp_all.d = abs(m_exp_all.d),
    partner_all.d = abs(m_partner_all.d)
  ) %>%
  mutate(across(where(is.numeric), inline_hook)) %>%
  unite(m_exp_all.conf.low, m_exp_all.conf.high, sep = ", ",
        col = "exp_all.conf") %>% 
  unite(m_partner_all.conf.low, m_partner_all.conf.high, sep = ", ",
        col = "partner_all.conf") %>% 
  mutate(
    condition = seq(0, 8, by = 2),
    exp_all.conf = paste0("[", exp_all.conf, "]"),
    partner_all.conf = paste0("[", partner_all.conf, "]")
    ) %>%
  select(condition, 
         m_exp_all.mean_0, m_exp_all.sd_0,
         m_exp_all.mean_1, m_exp_all.sd_1, 
         exp_all.conf, exp_all.d,
         m_partner_all.mean_0, m_partner_all.sd_0,
         m_partner_all.mean_1, m_partner_all.sd_1, 
         partner_all.conf, partner_all.d) %>%
  column_to_rownames("condition") %>%
  kable(
    digits = 2, format = "latex", booktabs = TRUE,
    caption = "\\textbf{Short-term mating behaviors of heterosexual men and gay men for 0-8 long-term relationship likelihood when gender differences existed in mating preferences}",
    row.names = TRUE, 
    col.names = c(rep(c(rep(c("M", "SD"), 2), "Mean difference 95 CI [LL, UL]","d"), 2))
    ) %>%
  row_spec(0, italic = TRUE, align = "c") %>%
  add_header_above(
    c(" " = 1,
      "Heterosexual men" = 2, "Gay men" = 2, " " = 2,
      "Heterosexual men" = 2, "Gay men" = 2, " " = 2)
    ) %>%
  add_header_above(
    c(" " = 1,
      "Average short-term mating experiences" = 6,
      "Average short-term mates" = 6)
    ) %>%
  column_spec(1, width = "0.5in") %>%
  footnote(
    general = "\\\\textit{Note.} \\\\textit{M} = Mean. \\\\textit{SD} = Standard deviation. CI = Confidence interval. LL = Lower limit. UL = Upper limit. The means were calculated using the full population of heterosexual men and gay men in the model.",
    general_title = "", escape = FALSE,
    threeparttable = TRUE, fixed_small_size = TRUE) %>% 
  kable_styling(latex_options = "scale_down")
```

\begingroup
\renewcommand{\arraystretch}{1.5}

\renewcommand{\thetable}{C3}

\begin{table}

\caption{\textbf{Short-term mating behaviors of heterosexual males and gay males for 0-8\% long-term relationship likelihood when sex differences existed in mating preferences}}
\centering
\begin{threeparttable}
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{0.5in}cccccccccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{6}{c}{Average short-term mating experiences} & \multicolumn{6}{c}{Average short-term mates} \\
\cmidrule(l{3pt}r{3pt}){2-7} \cmidrule(l{3pt}r{3pt}){8-13}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{\makecell[c]{Heterosexual\\males}} & \multicolumn{2}{c}{Gay males} & \multicolumn{2}{c}{ } & \multicolumn{2}{c}{\makecell[c]{Heterosexual\\males}} & \multicolumn{2}{c}{Gay males} & \multicolumn{2}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11}
\multicolumn{1}{c}{\makecell[c]{Long-term\\Likelihood\\(\%)}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\makecell[c]{Mean difference\\95\%CI [LL, UL]}} & \multicolumn{1}{c}{\makecell[c]{Cohen's \\ \em{d}}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\makecell[c]{Mean difference\\95\%CI [LL, UL]}} & \multicolumn{1}{c}{\makecell[c]{Cohen's \\ \em{d}}}\\
\midrule
\multicolumn{1}{c}{0} & 5.94 & 0.61 & 8.75 & 0.76 & {}[–3.00, –2.62] & 4.06 & 3.43 & 0.34 & 5.42 & 0.43 & {}[–2.10, –1.89] & 5.15\\
\multicolumn{1}{c}{2} & 2.57 & 0.30 & 5.02 & 0.48 & {}[–2.57, –2.34] & 6.17 & 1.65 & 0.17 & 3.38 & 0.30 & {}[–1.80, –1.66] & 6.99\\
\multicolumn{1}{c}{4} & 1.97 & 0.29 & 3.92 & 0.47 & {}[–2.06, –1.84] & 4.94 & 1.23 & 0.15 & 2.67 & 0.28 & {}[–1.50, –1.38] & 6.35\\
\multicolumn{1}{c}{6} & 1.66 & 0.27 & 3.22 & 0.43 & {}[–1.67, –1.46] & 4.34 & 1.02 & 0.13 & 2.21 & 0.26 & {}[–1.25, –1.14] & 5.85\\
\multicolumn{1}{c}{8} & 1.53 & 0.26 & 2.88 & 0.39 & {}[–1.44, –1.25] & 4.01 & 0.88 & 0.12 & 1.92 & 0.22 & {}[–1.08, –0.99] & 5.93\\
\bottomrule
\end{tabular}}
\begin{tablenotes}
\small
\item \textit{Note.} \textit{M} = Mean. \textit{SD} = Standard deviation. CI = Confidence interval. LL = Lower limit. UL = Upper limit. The means were calculated using the full population of heterosexual males and gay males in the model (\textit{N\textsubscript{hetero}} = \textit{N\textsubscript{gay}} = 150).
\end{tablenotes}
\end{threeparttable}
\end{table}

\endgroup


\elandscape

## References

<div id="refs"></div>
